Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2020-09-17 manuscript/manuscript.Rmd
Suparna edits

---
title: "Who Cares About Crackdowns? Exploring Individual Donor Responses to Closing Civic Space Abroad"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 5, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past decade, international nonprofit organizations (INGOs) have faced increasing legal scrutiny and closed civic space abroad. While we know this has negatively impacted public funding flows, we know little about how this has affected private funding flows, especially be individuals. This turn to private aid introduces potentially new dynamics in INGO fundraising strategies, since individuals have a range of demographic characteristics and past experiences with philanthropy that shape their propensity to donate. Which demographic and experiential factors influence the decision to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017]. Consequently, official aid flows to nongovernmental organizations (NGOs) have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know that there has been a negative impact on funding flows from public sector organizations like USAID, DFID, and others, we know relatively little about how this global crackdown affects private funding flows, in particular from individual donors.

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is quite robust across a variety of contexts. But the dynamics of *international* giving are under-examined in the literature on individual donor preferences.[^exceptions]. Research on the determinants of philanthropy towards international causes—particularly for organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Understanding why individuals donate to international causes is especially essential to understand, given that such philanthropy can be harder to motivate because the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100].

The amount of money donated by private individuals to CSOs working abroad, however, has steadily increased over time. 
Total aggregate philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, total private giving to CSOs working specifically in international affairs reached $22.08 billion. More importantly, organizations working in the area of international affairs are one of the only two types of recipient organizations that have continued experiencing substantial growth in giving—even at a time when giving to religious institutions, education, human services and arts, culture and society has been on the decline [@givingusa:2019].[^intlaffairs]

@@

While this may be a cause for optimism among CSOs working internationally, we know little about the individuals who donate to CSOs working in other countries. We also lack any systematic understanding of how such giving is affected by an increasingly restricted civic space abroad. Given these challenges, what is the optimal mix of individual-level donor attributes for international NGOs working in restricted political environments interested in maximizing donations? What kinds of individual donors are more or less likely to consider a CSO's host country political environment, its transparency and accountability practices, and its funding structure when making the decision to donate? How do individual-level donor characteristics such as political ideology, knowledge, activism and other demographic characteristics interact with organizational-level attributes to influence philanthropic decisions?

We explore these questions using a simulated market for philanthropic donations, based on data and models from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. In particular, personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism tend to prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments and prefer instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits and charitable giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and prior experiences that make certain individuals more likely to donate can be very important to nonprofits, as this information can help them craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most CSOs and NGOs working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to private individuals.

Framing is particularly important in this regard—how people process information affects their willingness to donate.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple factors, it can help these CSOs frame their appeals.

More broadly, while research in public policy and nonprofit studies has focused on charitable giving, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international nongovernmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

Below, we first explore how the global crackdown on civil society has led many CSOs working internationally to shift their fundraising strategies towards individual private donors rather than more restricted government and foundation sources. We then examine the state of the existing literature on the key drivers of individual charitable giving, as understanding what motivates private donations is crucial to more effective CSO fundraising. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. One primary reason for this turn to civil society is that many government and foundation donors perceive that NGOs have a superior ability to achieve intended outcomes. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis. More broadly, a robust civil society is also considered to be one of the necessary preconditions for a healthy, functioning democracy due to its ability to generate social capital and hold governments accountable [@Putnam:1995].

## Sources of foreign aid

Thousands of international NGOs and CSOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. In this paper, we are most interested in individuals as a funding source, and we examine individual-level determinants to foreign giving. This is important to investigate as individual private donors may not necessarily mirror public donor choices. Surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, foreign aid and democracy aid is often considered unresponsive to developments on the ground [@brainard2007; @Carothers:2015] and public aid decisions—especially on final recipients— are not easily accessible. There is also substantial variety in how donor governments allocate funds. Emerging donor governments differ from established donors in that they demand fewer human rights, governance or environmental conditions preferred by Western donors [@fenglerkharas].

With the advent of funding restrictions on NGOs globally and a subsequent decline in official government aid and foundation funding, many organizations are being forced to individual donors abroad. Donations to international affairs organizations has steadily increased over the past two decades, with a noticeable rise in the past decade. Between 2014 and 2016, giving to international affairs increased by 14.1% [@givingusa2017report]. Most of this is driven by individuals making small-scale donations. In 2017, individuals gave \$286.7 billion, or approximately 70% of total giving to organizations working in international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [givingusa:2019]. Given that a majority of this research is restricted to elite, high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature in charitable giving looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, the decision to donate to an NGO is not determined solely by an organization's characteristics. Donors have particular personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies examines the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature further concentrates on a specific kind of social capital—associational capital, or "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not) were both more likely to make charitable donations and make bigger donations. This is because volunteering exposes individuals to the need for donations to charities and nonprofits and made them feel more compelled to do so [@hossainlamb2012].

Higher education can also play an important role in charitable giving decisions as such education fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. This in turn has been shown to be particularly vital in giving to international causes. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. An additional year of education has been shown to increase the likelihood of giving to international causes by about 2 percentage points [@OktenOsili:2007].

Higher education also helps determine the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—all issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions. However, religiosity does not necessarily have the same effect on giving to all causes. @neumayrhandy2019 find that religiosity is negatively associated with giving to environmental issues and animal welfare, while having an insignificant effect on giving to international relief.

Finally, recent research has also looked at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad. More specifically, research finds that immigrant households are more likely to contribute to international aid organizations [@OktenOsili:2007], and that immigrants can find value in donating to both home and host cultures, especially if family members are still residing in the home country [@burgoyne2005].

Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. While we expect these determinants to be reflected in our survey experiment, we do not lay out specific hypotheses. Because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Our innovative experimental approach, however, provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. We then displayed various combinations of features, including one of four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), one of two organizational practices (financial transparency and third-party accountability audits), one of three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and one of three relationships with host governments (friendly, criticized by the government, and under government crackdown). We presented twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination.

In June 2019, We recruited survey respondents through Centiment, a commercial provider of high quality non-probability opt-in survey panels. Sentiment ensures panel quality by actively recruiting representative samples of the US population and rewards participants with monetary incentives. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

@@

In previous work [@ChaudhryDotsonHeiss:2019], we use the results of this survey experiment to measure how the organizational characteristics of an NGO can influence donor preferences. We find that donors prefer to give money to organizations that engage in transparency and accountability practices, receive funding from many individual donors, and that work on humanitarian issues like disaster and refugee relief. Importantly, we also find that donors are less likely to be interested in donating to NGOs facing government crackdown abroad. These results provide unique insight for NGOs looking to raise money from private donors and can influence how these organizations might frame their calls for fundraising.

However, the findings in @ChaudhryDotsonHeiss:2019 focus only on organizational attributes and do not distinguish between different types and specific characteristics of donors. As part of the survey experiment, we collected information about the survey respondents themselves, including their demographic characteristics and their interest and involvement in public affairs, charity, voluntarism, politics, and religion.

In this paper, we focus instead on exploring which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics on every combination of organizational characteristics is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. In our previous work—and in most research that relies on regression model—the analytic focus is on individual regression coefficients and the effect of single covariates of interest. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously

We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes). This combination of personas and organizations yields 768 hypothetical scenarios where potential donors can interact with different organizations.

@@

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, most do not. This is acceptable in simulations, and each combination of persona characteristics is plausible in the real world. Our analysis uses the characteristics of our sample to model the effects of various salient attributes while holding others constant, and thus allows us to explore multiple attributes at once. Also importantly, our findings are not causal. Rather, they are the descriptive results of simulated data from a nationally representative sample of people who regularly donate to charity. The insights from this simulation, however, still reveal interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad.

@@

Figures @fig:issue-social, @fig:relationship-social, and @fig:funding-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue. Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

Figure @fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.

>>>>

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 8, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international nonprofit organizations (INGOs) have faced increasing legal scrutiny abroad, which has negatively impacted public funding flows. Consequently, INGOs may feel the need to fundraise from individuals. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017].[^NGOdef] As part of this phenomenon, CSOs, particularly NGOs, face increasing obstacles to their work, including legal restrictions and threats to their freedom of association, expression and speech. Legal restrictions-a core part of these efforts-create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct or repress these organizations-what we refer to as legal crackdowns in this paper.

As a result of these legal crackdowns, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like USAID, DFID, and other OECD funding agencies, philanthropy from private donors, in particular, foundations, appears to not be as adversely affected [@McGill:2018]. Given this trend of closing civic space, NGOs need access to multiple sources of funding and individual-level giving can serve as one such source of funds. What individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy towards their operations abroad?

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is robust across a variety of contexts. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Research on the determinants of philanthropy towards international causes—particularly for organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Understanding why individuals donate to international causes is essential as it can be harder to motivate as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100].

Individual giving is also important to understand because the amount of money donated by private individuals to CSOs working abroad has steadily increased over time. Total aggregate philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, total private giving to CSOs working specifically in international affairs reached \$22.08 billion. Moreover, organizations working in the area of international affairs are one of the only two types of recipient organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this is driven by individuals making small-scale or micro-donations.[^small donors] In 2017, individuals gave \$286.7 billion, or approximately 70% of total giving to organizations working in international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. However, we lack a systematic understanding of which individual donors may be further motivated or deterred to donate under an increasingly restricted civic space abroad. What is the optimal mix of individual-level donor attributes for international NGOs working in restricted political environments interested in maximizing donations? What kinds of individual donors are more or less likely to consider an NGO's host country political environment when making the decision to donate?

@@

We explore these questions using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments, preferring instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits and charitable-giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational-level factors, it can help these CSOs frame their appeals.

More broadly, while research in public policy and nonprofit studies has focused on giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international nongovernmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

Below, we first examine the state of the existing literature on the key drivers of individual charitable giving, as understanding what motivates private donations is crucial to more effective NGO fundraising. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive that NGOs have a superior ability to achieve intended outcomes. Therefore, in countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis.

## Sources of funding

Thousands of INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles especially from a funding perspective. Official aid donors have reduced the amount of aid channeled to countries with legal restrictions @DupuyPrakash:2018]. In addition, official donors have also redirected funds within restrictive countries, decreasing funds for politically sensitive causes in favor of tamer issues that are more compatible with the governments of other countries [@ChaudhryHeiss:2018]. However, it is notable that private funding towards international nonprofits has not been as adversely affected. Many foundations have continued channeling funds to countries with hostile legal environments [@FoundationCenter2018].

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Between 2014-2016, giving to organizations working in international affairs increased by 14.1% [@givingusa2017report]. Moreover, at least in the U.S. context, donations from individuals far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion from corporations [Source from NVSQ]. Given that a majority of research is restricted to elite high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well. In this paper, we therefore look at which kinds of individual donors—beyond high-net worth donors—are most likely to give abroad.[^highnetworth]

@@

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, donation decisions are not determined solely by an organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local—rather than international—causes. This literature focuses on the importance of five main attributes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies examines the role of social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Research finds that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not) were both more likely to make charitable donations and make bigger donations. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012].

Higher education can also play an important role in charitable-giving decisions as it fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations, which is vital in giving to international causes [@BekkersWiepking:2011]. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

Higher education also helps determine the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions.

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad. More specifically, immigrant households are more likely to contribute to international aid organizations [@OktenOsili:2007], and that immigrants can find value in donating to both home and host cultures, especially if their family members still reside in the home country [@burgoyne2005].

Overall, individuals with higher levels of income, education, and greater religious proclivities are not only more likely to give internationally, but also give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. While we expect these determinants to be reflected in our survey experiment, we do not lay out specific hypotheses. Because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Our innovative experimental approach, however, provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. These organizations are commonly used in surveys on TODO TODO, in addition to other organizations such as MSF, World Vision (LOOK UP SAVED PAPERS - I THINK NVSQ). We then displayed various combinations of features, including one of four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), one of two organizational practices (financial transparency and third-party accountability audits), one of three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and one of three relationships with host governments (friendly, criticized by the government, and under government crackdown). We presented twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination.

In June 2019, we recruited survey respondents through Centiment, a commercial provider of high quality non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and rewards participants with monetary incentives. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously

We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, most do not. This is acceptable in simulations, and each combination of persona characteristics is plausible in the real world. Our analysis uses the characteristics of our sample to model the effects of various salient attributes while holding others constant, and thus allows us to explore multiple attributes at once.

Also importantly, our findings are not causal. Rather, they are the descriptive results of simulated data from a nationally representative sample of people who regularly donate to charity. The insights from this simulation, however, still reveal interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad.

@@

@fig:issue-social, @fig:relationship-social, and @fig:funding-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue.

Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

@fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.
